---
title: "Honey Bees - edgeR-btwnLane-QLF Interaction"
author: Lindsay Rutter
output:
  packagedocs::package_docs:
    toc: true
    toc_collapse: true
vignette: >
  %\VignetteEngine{packagedocs::redirect}
---
  
<meta http-equiv="content-type" content="text/html;charset=utf-8" />
  
```{r global_options, include=FALSE}
# R output pre blocks are styled by default to indicate output
knitr::opts_chunk$set(
  comment = NA,
  cache = TRUE,
  fig.height = 8,
  fig.width = 10
)
```

```{r}
library(edgeR)
library(ggplot2)
library(GGally)
library(EDASeq)
library(utils)

thisPath <- getwd()

beeCounts <- readRDS("../../../data/data.Rds")
beeCounts <- as.matrix(beeCounts)

y <- DGEList(counts=beeCounts)
```

edgeR vignette states that a gene is required to have a count of 5-10 in a library to be considered expressed in that library. Here minLib is equal to `r min(y$samples$lib.size)`. A CPM of 4 corresponds to a count of ~1 in the minimum number of samples in a group (6). So, I keep only rows that have that. This reduces the number of genes from 15,314 to 8,672.

```{r}
minLib <- min(y$samples$lib.size)
keep <- rowSums(cpm(y)>4) >= 6
# Number of genes 15,314--> 8,672
y <- y[keep, , keep.lib.sizes=FALSE]

# Next, I used edgeR normalization.
y[[1]] <- betweenLaneNormalization(y[[1]], which="full", round=FALSE)
y$samples$group <- unlist(lapply(colnames(y), function (x) unlist(strsplit(x, "[.]"))[1]))
Group = factor(y$samples$group)
#targets <- data.frame(rownames=paste0("Sample",1:24), Virus = unlist(lapply(colnames(y), function (x) substring(unlist(strsplit(x, "[.]"))[1],1,1))), Diet = unlist(lapply(colnames(countdata), function (x) substring(unlist(strsplit(x, "[.]"))[1],2,2))), Group = Group)
```

```{r}
# Follow edgeR vignette (3.2.4)
design <- model.matrix(~0+Group, data=y$samples)
colnames(design) <- levels(Group)
y <- estimateDisp(y, design)
fit <- glmQLFit(y, design)
```

```{r}
my.contrasts <- makeContrasts(
  NC.NR = NC-NR,
  NC.VC = NC-VC,
  NC.VR = NC-VR,
  NR.VC = NR-VC,
  NR.VR = NR-VR,
  VC.VR = VC-VR,
  NR.NC = NR-NC,
  VC.NC = VC-NC,
  VR.NC = VR-NC,
  VC.NR = VC-NR,
  VR.NR = VR-NR,
  VR.VC = VR-VC,
  aveC.aveR = (NC+VC)/2-(NR+VR)/2,
  aveN.aveV = (NC+NR)/2-(VC+VR)/2,
  aveR.aveC = (NR+VR)/2-(NC+VC)/2,
  aveV.aveN = (VC+VR)/2-(NC+NR)/2,
  intN.V.C.R = (NC-NR)-(VC-VR),
  intN.V.R.C = (NR-NC)-(VR-VC),
  intC.R.N.V = (NC-VC)-(NR-VR),
  intC.R.V.N = (VC-NC)-(VR-NR),
  intV.N.C.R = (VC-VR)-(NC-NR),
  intV.N.R.C = (VR-VC)-(NR-NC),
  intR.C.N.V = (NR-VR)-(NC-VC),
  intR.C.V.N = (VR-NR)-(VC-NC),
levels=design)
```

Below we save the DEGs from all pairwise combinations of treatment groups.

```{r}
allPairs = data.frame(Combination = character(), NumberDEG = numeric())
dataMetrics <- list()
contNames = attributes(my.contrasts)$dimnames$Contrasts

for (i in 1:(length(contNames))){
  lrt <- glmQLFTest(fit, contrast=my.contrasts[,contNames[i]])
  lrt <- topTags(lrt, n = nrow(y[[1]]))[[1]]
  lrt5 <- lrt[which(lrt$FDR<0.05),]
  
  lrtLength <- nrow(lrt5)
  allPairs = rbind(allPairs, data.frame(Combination = contNames[i], NumberDEG = lrtLength))
  
  metrics = cbind(ID = rownames(lrt), lrt)
  metrics$ID = as.character(metrics$ID)
  dataMetrics[[contNames[i]]] <- metrics
}
allPairs <- allPairs[order(allPairs$NumberDEG),]
saveRDS(dataMetrics, file="dataContrasts.Rds")
```

```{r}
allPairs
```